---
title: "DS 202 Final Project Data Dynamos"
author: "Logan Becker, Gavin Herum, Jackson Weaver"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## View Data Set

```{r, message = FALSE}
#install.packages("readr")
library(readr)
# read datasets
injury_data <- read_csv("./injury_data.zip") 
player_data <- read_csv("./player_data.csv")
seasonsStats_data <- read_csv("./Seasons_Stats.csv")
head(injury_data)
head(player_data)
head(seasonsStats_data)
```

## Filter every dataset to be all years beyond (inclusive) 2010

```{r, message=FALSE}
# Filter injury_data
filtered_injurydata <- injury_data[as.numeric(format(injury_data$Date, "%Y")) >= 2010,]
print(paste('injury_data'))
summary(filtered_injurydata["Date"])

# Filter player_data
filtered_playerdata <- subset(player_data, year_start >= 2010 | year_end >= 2010)
print('player_data')
summary(filtered_playerdata[c("year_start", "year_end")])

# Filter seasons_stats
filtered_seasonsStatsdata <- subset(seasonsStats_data, Year >= 2010)
print('seasonsStats_data')
summary(filtered_seasonsStatsdata["Year"])
```

## Change old team names to their new team name

```{r, message=FALSE}
all_teams <- unique(filtered_injurydata$Team)
all_teams
# We see there is 32 teams here, there should only be 30 nba teams. Bullets and Bobcats are old names, let's change those to be correct
# Hornets <- Bobcats
# Wizards <- Bullets

filtered_injurydata$Team <- 
  ifelse(filtered_injurydata$Team == "Bobcats", "Hornets",
  ifelse(filtered_injurydata$Team == "Bullets", "Wizards",
         filtered_injurydata$Team))
# Check if there is now 30 teams
all_teams <- unique(filtered_injurydata$Team)
all_teams
```

## Lets find out the average duration of an injury

The average duration of an injury in the NBA is important to know because it helps make informed decisions on recovery and what to expect for a player. While the average duration of an injury cannot inform any specific recovery. It can help us get an understanding of the expected results of an injury. Knowing how long an injury could last informs teams, players, and spectators of a sport.

```{r, message = FALSE}

library(magrittr)
library(dplyr)
library(ggplot2)

# Create subset of the end of an injury
Acquired_subset <- subset(filtered_injurydata, is.na(filtered_injurydata$Relinquished))
Acquired_subset <- subset( Acquired_subset, select = -c(1, Relinquished) )

# Create subset of the begining of an injury
Relinquished_subset <- subset(filtered_injurydata, is.na(filtered_injurydata$Acquired))
Relinquished_subset <- subset( Relinquished_subset, select = -c(1, Acquired) )


# Merge the dataset with itself based on the player's name
merged_data <- merge(x = Acquired_subset, y = Relinquished_subset, by.x = "Acquired", by.y = "Relinquished", suffixes = c("_end", "_begin"))

merged_data$Date_gap <- as.numeric(difftime(merged_data$Date_end, merged_data$Date_begin, units = "days"))

# Removing all Date_gaps less than 0
merged_data <- subset(merged_data, Date_gap > 0 )

# Filtering to ensure no duplicate Date_end for a given Date_begin
injury_duration <- merged_data %>% 
  group_by(Acquired, Date_end)  %>% 
  filter(Date_gap == min(Date_gap))

# Reordering collumns of data
injury_duration <- injury_duration[, c("Acquired", "Date_begin", "Date_end", "Date_gap", "Team_begin", "Team_end", "Notes_begin", "Notes_end")]

# Rename column to Player
colnames(injury_duration)[colnames(injury_duration) == "Acquired"] <- 'Player'

Injury_duration_average <- mean(injury_duration$Date_gap)
Injury_duration_average

filtered_injury_duration <- subset(injury_duration, Date_gap <= 365 )
Injury_duration_average <- mean(filtered_injury_duration$Date_gap)
Injury_duration_average

# This graph is messed up because Patty Mills had an injury that lasted for 2976 days, which is not accurate to reality
ggplot(injury_duration, aes(x = Date_gap)) +
  geom_bar(fill = "skyblue", color = "black") +
  labs(title = "Frequency of Different Injury Durations",
       x = "Duration of injury",
       y = "Frequency") +
  theme_minimal()

# This graph is a zoomed in version of only the injuries that lasted less than a year, this should filter out the injuries that were not accurate.
ggplot(filtered_injury_duration, aes(x = Date_gap)) +
  geom_bar(fill = "skyblue", color = "black") +
  labs(title = 
  "Frequency of Different Injury Durations
  Excluding Injuries that Lasted more than a year (365 days)",
       x = "Duration of injury",
       y = "Frequency") +
  theme_minimal()
```

Based off of this data the average duration of an injury will be 12.24041 days when you exclude the extereme injuries that lasted more than a year. We can see that the data is massively skewed towards shorter injuries. The reason for this being that many of the data points aren't injuries at all and instead are illness. Some instants of players being placed on IL don't have any detail of why. This makes it hard for us to filter out only injuries.

## Injury per Player data setup

```{r, message=FALSE}
library(dplyr)
library(lubridate)
#change injurydata date column to just the year instead of date format to help merge
changed_injurydata <- filtered_injurydata
changed_injurydata$Date <- as.Date(changed_injurydata$Date)
changed_injurydata$Date <- year(changed_injurydata$Date)

# merge datasets and combine player info and injured player
ageVSinjury <- inner_join(changed_injurydata, filtered_seasonsStatsdata, 
                         by = c("Relinquished" = "Player", "Date" = "Year"),
                         relationship = "many-to-many")

# make sure there are no repeats (distinct IDs)
ageVSinjury <- distinct(ageVSinjury,...1.x, .keep_all = TRUE)
```

## Injuries per Player (2010 - 2017)

```{r, message=FALSE}
# First look at age range
age_range <- summary(filtered_seasonsStatsdata$Age)
age_range 
# Age range is 19-40, lets split into groups by 3-4

# With this knowledge we can make a function to split up a dataset into a list with different age groups
SplitAges <- function(df) {
  age1 <- subset(df, Age <= 22)
  age2 <- subset(df, Age <= 26 & Age > 22)
  age3 <- subset(df, Age <= 30 & Age > 26)
  age4 <- subset(df, Age <= 35 & Age > 30)
  age5 <- subset(df, Age > 35)
  return(list(age1 = age1,age2 = age2,age3 = age3,age4 = age4,age5 = age5)) #Returns list of 5 lists of age groups
}
# Split age range by calling function
ageInjury_list <- SplitAges(ageVSinjury["Age"])
agePlayer_list <- SplitAges(filtered_seasonsStatsdata["Age"])

# Function to print/save age ratios for each group
AgeRatios <- function(list1,list2,PRINT) {
  if(PRINT == TRUE) {
    print(paste("19-22 years ratio:", nrow(list1[[1]]) / nrow(list2[[1]])))
    print(paste("23-26 years ratio:", nrow(list1[[2]]) / nrow(list2[[2]])))
    print(paste("27-30 years ratio:", nrow(list1[[3]]) / nrow(list2[[3]])))
    print(paste("31-35 years ratio:", nrow(list1[[4]]) / nrow(list2[[4]])))
    print(paste("36-40 years ratio:", nrow(list1[[5]]) / nrow(list2[[5]])))
  } else {
    ratio1 <- nrow(list1[[1]]) / nrow(list2[[1]])
    ratio2 <- nrow(list1[[2]]) / nrow(list2[[2]])
    ratio3 <- nrow(list1[[3]]) / nrow(list2[[3]])
    ratio4 <- nrow(list1[[4]]) / nrow(list2[[4]])
    ratio5 <- nrow(list1[[5]]) / nrow(list2[[5]])
    return(list("19-22 years" = ratio1,"23-26 years" = ratio2,"27-30 years" = ratio3,"31-35 years" = ratio4,"36-40 years" = ratio5))
  }
}

# Print out ratios of injured players vs all players for each age range
AgeRatios(ageInjury_list, agePlayer_list, TRUE) # PRINTS AGE RATIOS
listAgeRatios <- AgeRatios(ageInjury_list, agePlayer_list, FALSE) # SAVES AGE RATIOS AS LIST
```

## Graph for Injuries per player

```{r, message=FALSE}
library(ggplot2)
# Convert ratios to data frame
dfAgeRatio <- data.frame(Age_Group = names(listAgeRatios), Ratio = unlist(listAgeRatios), row.names = NULL)
dfAgeRatio
# Create graph to display Age vs Injury data
ggplot(dfAgeRatio, aes(x = Age_Group, y = Ratio, fill = Age_Group)) +
  geom_bar(stat = "identity") +
  labs(title = "Injuries per Player by Age Group in the NBA", x = "Age Group", y = "Injuries per Player") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Statistics for injury duration by age group

```{r, message=FALSE}
library(lubridate)
# Change date column to just year
changed_injuryduration <- injury_duration
changed_injuryduration$Date_begin <- as.Date(changed_injuryduration$Date_begin)
changed_injuryduration$Date_begin <- year(changed_injuryduration$Date_begin)

# Merge injury_duration and filtered_seasonsStatsdata
durationANDage <- inner_join(changed_injuryduration, filtered_seasonsStatsdata, 
                             by = c("Date_begin" = "Year", "Player" = "Player"),
                             relationship = "many-to-many")
# Split by age group
ageInjuryDuration_list <- SplitAges(durationANDage[c("Date_gap", "Age")])

# Statistics for each age group
statisticCalculations <- function(df, fun) {
  fun1 <- fun(df[[1]]$Date_gap)
  fun2 <- fun(df[[2]]$Date_gap)
  fun3 <- fun(df[[3]]$Date_gap)
  fun4 <- fun(df[[4]]$Date_gap)
  fun5 <- fun(df[[5]]$Date_gap)
  return(list("19-22 years" = fun1, "23-26 years" = fun2, "27-30 years" = fun3, "31-35 years" = fun4, "36-40" = fun5))
}
# Get lists for each age group for means, maxs, medians, sds
meanDurations <- statisticCalculations(ageInjuryDuration_list, mean)
maxDurations <- statisticCalculations(ageInjuryDuration_list, max)
medianDurations <- statisticCalculations(ageInjuryDuration_list, median)
sdDurations <- statisticCalculations(ageInjuryDuration_list, sd)
```

## Graph Injury duration by age group

```{r, message=FALSE}
#install.packages('cowplot')
library(cowplot)
library(ggplot2)
# Create datasets for the statistics for graph building
mean_df <- data.frame(Age_Group = names(meanDurations), Mean = unlist(meanDurations))
max_df <- data.frame(Age_Group = names(maxDurations), Max = unlist(maxDurations))
median_df <- data.frame(Age_Group = names(medianDurations), Median = unlist(medianDurations))
sd_df <- data.frame(Age_Group = names(sdDurations), SD = unlist(sdDurations))


statistic_barplot <- function(data, x, y, fill, title, y_label) {
  plot <- ggplot(data, aes_string(x = x, y = y, fill = fill)) +
    geom_bar(stat = "identity") +
    labs(title = title, x = "Age Group", y = y_label) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
          plot.title = element_text(size = 11),
          axis.title = element_text(size = 8),
          legend.text=element_text(size=8),
          legend.title = element_text(size = 10))
  return(plot)
}
# Create barcharts for Mean, Max, Median,  by age group
bar_means <- statistic_barplot(mean_df, "Age_Group", "Mean", "Age_Group", "Mean Injury Duration by Age Group", "Mean Injury Duration (Days)")
bar_maxs <- statistic_barplot(max_df, "Age_Group", "Max", "Age_Group", "Longest Injury Duration by Age Group", "Longest Injury Duration (Days)")
bar_medians <- statistic_barplot(median_df, "Age_Group", "Median", "Age_Group", "Median Injury Duration by Age Group", "Median Injury Duration (Days)")
bar_sd <- statistic_barplot(sd_df, "Age_Group", "SD", "Age_Group", "Standard Deviation Injury Duration by Age Group", "Standard Deviation Injury Duration (Days)")

plot_grid(bar_means, bar_medians, bar_maxs, bar_sd, ncol = 2, nrow = 2, align = "v")

```

## may not be needed but here for testing purposes and to safe keep

```{r, message = FALSE}
# Filter injury data to only have rows where Relinquished is not NA (someone is injured)
library(dplyr)
filtered_injurydata <- filtered_injurydata %>%
  filter(!is.na(Relinquished))
# Deletes acquired row
#filtered_injurydata <- subset(filtered_injurydata, select = -Acquired)
summary(filtered_injurydata)


# Filter injury data to include only notes that contain "IL,IR,DTD" (injuries)
filtered_injurydata <- subset(filtered_injurydata, grepl("IL|IR|DTD", Notes))
print(filtered_injurydata)
```

## Let's look at the highest injured teams

```{r, message=FALSE}
# MIGHT NOT KEEP THIS CODE WHEN PRESENTING UNLESS WE WANT TO LOOK AT SOMETHING FURTHER WITH IT
team_counts <- table(filtered_injurydata$Team)
sorted_team_counts <- sort(team_counts, decreasing = TRUE)
highest5_injuredTeams <- head(sorted_team_counts, 5)
highest5_injuredTeams
```

The most frequently injured team are Spurs (582), Celtics (530), Raptors (529), Mavericks (497), Heat (474). \>\>\>\>\>\>\> 6bfaeee75f4c80ebdbb9a8326c055675cb8c7f13

```{r}

#combining data from injury_data and player_data, replacing values in Acquired column if NA shows up in Acquired column with Relinquished value in row. Then changing name to equal Acquired.

#We are making a column to merge by in the future

combining_data <- injury_data %>% select(Acquired, Relinquished, Notes) %>% mutate(Acquired = if_else(is.na(Relinquished), Acquired, Relinquished)) %>% mutate(name = Acquired) 


#made a new column with name to align with player_data, but no distinction among names, so same name will show up multiple times

#below will merge the two dataframes by name

merged_playerData_combiningData <- merge(player_data,combining_data, by = "name")

#count of injuries per player
counting_number_injuries <- combining_data %>% count(Acquired) #%>% as.numeric(n) 

ggplot(counting_number_injuries, aes(x = n)) +
  geom_histogram(binwidth = 1,) +
  labs(title = "Frequency of Injuries", x = "number of injuries", y = "Frequency")

#counts the number of injuries 
#counting_number_of_injuries <- counting_number_injuries %>% rename(number_of_injuries = n) %>% count(number_of_injuries) %>% rename(people_injured = number_of_injuries) %>% rename(count_of_injuries = n)

#print(counting_number_of_injuries)

merged_playerData_combiningData <- merge(player_data,combining_data, by = "name")




# shows number of injuries based on position
injuries_by_position <- merged_playerData_combiningData %>% count(position) %>% rename(number_of_injuries = n)

library(ggplot2)

#creates bar graph of injuries by position

ggplot(injuries_by_position, aes(x = position, y = number_of_injuries)) +
  geom_bar(stat = "identity") +
  labs(x = "Position Injuries", y = "Frequency", title = "Frequency of Injuries by Position")

##IL means they won't take up a bench spot
##IR means they'll be taking up a bench spot

##would certain positions get put on IL instead of IR because of position played?

library(stringr)
IL <- "placed on IL"
IR <- "placed on IR"

#below gives injuries that start with "placed on IL"

IL_position_injuries <- merged_playerData_combiningData %>% filter(str_starts(Notes, IL)) %>% select(position, Notes) %>% count(position) %>% rename(IL_injuries = n)

#below gives something that starts with IR and counts the position of each person

IR_position_injuries <- merged_playerData_combiningData %>% filter(str_starts(Notes, IR)) %>% select(position, Notes) %>% count(position) %>% rename(IR_injuries = n)
#print(IR_position_injuries)

#above from counting position for people placed on IR shows guard and forward are placed on it the most, could possibly be due to aggressiveness of position needed, however F-G, and Guard forward are the lowest

ggplot(IR_position_injuries, aes(x = position, y = IR_injuries)) +
  geom_bar(stat = "identity") +
  labs(x = "Position", y = "Frequency", title = "IR position injuries")

##how many IL happen per position, but by same position how many IR happens?

##IL means a more serious injury which could then impact that persons life. 
##Would they put someone on IL if they had a certain position compared to another position?

ggplot(IL_position_injuries, aes(x = position, y = IL_injuries)) +
  geom_bar(stat = "identity") +
  labs(x = "Position Injuries", y = "Frequency", title = "IL Position Injuries")


#merging IR and IL injuries in order to compare in bar graph
merging_IR_IL <- merge(IL_position_injuries, IR_position_injuries, by = "position")
#print(merging_IR_IL)


#bar graph below shows how much more likely you are to get put on IL when at that position.
#Doesn't account for type of injury, aggressiveness needed at said position and how they affect the team chemistry

ggplot(merging_IR_IL, aes(x = position, y = IL_injuries / IR_injuries)) +
  geom_bar(stat = "identity") +
  labs(x = "Position", y = "IL Injuries per IR Injuries", title = "IL to IR Injury constant Based on Position")


library(tidyr)


#below gets the number of players injured
injured_sample <- merged_playerData_combiningData %>% distinct(name) %>% nrow()

#below gets the total number of players
overall_sample_size <- player_data %>% distinct(name) %>% nrow()

#below gets the injury at least once per total players
print(injured_sample / overall_sample_size) * 100

#above shows we got a 46.5% chance a player will get injured at least once, while NLH says the amount injured at least once is a 46.8% chance
 

```
