---
title: "DS 202 Final Project Data Dynamos"
author: "Logan Becker, Gavin Herum, Jackson Weaver"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

We chose to look at NBA injury data. The data set we chose was from Kaggle. We mainly wanted to look into injury duration, injuries by position, and injuries by age group. The main goal of this report is to find and take note of interesting trends in NBA injury data.

# Data sets, cleaning, and filtering

## View Data Set

```{r, message = FALSE}
#install.packages("readr")
library(readr)
# read datasets
injury_data <- read_csv("./injury_data.zip") 
player_data <- read_csv("./player_data.csv")
seasonsStats_data <- read_csv("./Seasons_Stats.csv")
head(injury_data)
head(player_data)
head(seasonsStats_data)
```
## Data Structure
For the NBA injury stats we have 1 csv file with 6 columns (Index,Date,Team,Acquired,Relinquished,Notes). This csv file includes 37,667 rows. For the NBA Player stats we have 3 csv files. Two of the csv files are very similar and we will not need the Player csv, so we do not import that into our analysis nor do we use it. First csv Seasons_Stats has 53 columns which we will only use 5 of the columns (Index, Year, Player, Position, Age). It contains 24,691 rows. The second csv file has 8 columns but we will only use 4 of the columns (name, year_start, year_end, position). This dataset included 4,550 rows.

## Filter every dataset to be all years beyond (inclusive) 2010

Here we want to simplify our data down to the range that we want. This means that we are limiting the data to the years between 2010-2023 inclusively. In hindsight, we should have spent the time to put each data point into their respective seasons and working based off of the NBA season calendar. We initially did this simplification because it was quick and got us close to what we wanted.

```{r, message=FALSE}
# Filter injury_data
filtered_injurydata <- injury_data[as.numeric(format(injury_data$Date, "%Y")) >= 2010,]
print(paste('injury_data'))
summary(filtered_injurydata["Date"])

# Filter player_data
filtered_playerdata <- subset(player_data, year_start >= 2010 | year_end >= 2010)
print('player_data')
summary(filtered_playerdata[c("year_start", "year_end")])

# Filter seasons_stats
filtered_seasonsStatsdata <- subset(seasonsStats_data, Year >= 2010)
print('seasonsStats_data')
summary(filtered_seasonsStatsdata["Year"])
```

## Change old team names to their new team name

We found that in the filtered data there were still teams with names that have since been changed. We wanted to change these team names so they wouldn't cause issues later down the road.

```{r, message=FALSE}
all_teams <- unique(filtered_injurydata$Team)
all_teams
# We see there is 32 teams here, there should only be 30 nba teams. Bullets and Bobcats are old names, let's change those to be correct
# Hornets <- Bobcats
# Wizards <- Bullets

filtered_injurydata$Team <- 
  ifelse(filtered_injurydata$Team == "Bobcats", "Hornets",
  ifelse(filtered_injurydata$Team == "Bullets", "Wizards",
         filtered_injurydata$Team))
# Check if there is now 30 teams
all_teams <- unique(filtered_injurydata$Team)
all_teams
```

## Data Variables

### injury_data.csv
##### Index: Index
##### Date: The date the action occurred on
##### Team: Team the action occurred on
##### Acquired: Player name the team acquired (recovering player or replacement player)
##### Relinquished: Player name the team relinquishes (injury or other leave)
##### Notes: Short note on the action that occurred

### Seasons_Stats.csv
##### Index: Index
##### Year: year of the season
##### Player: First and last name of the player
##### Pos: Position of the player
##### Age: Age of the player

### player_data.csv
##### name: First and last name of the player
##### year_start: Year the player started his career
##### year_end: Year the player ended his career
##### position: Position of the player

# Analysis

## Lets find out the average duration of an injury

The average duration of an injury in the NBA is important to know because it helps make informed decisions on recovery and what to expect for a player. While the average duration of an injury cannot inform any specific recovery. It can help us get an understanding of the expected results of an injury. Knowing how long an injury could last informs teams, players, and spectators of a sport.

```{r, message = FALSE}

library(magrittr)
library(dplyr)
library(ggplot2)

# Create subset of the end of an injury
Acquired_subset <- subset(filtered_injurydata, is.na(filtered_injurydata$Relinquished))
Acquired_subset <- subset( Acquired_subset, select = -c(1, Relinquished) )

# Create subset of the begining of an injury
Relinquished_subset <- subset(filtered_injurydata, is.na(filtered_injurydata$Acquired))
Relinquished_subset <- subset( Relinquished_subset, select = -c(1, Acquired) )


# Merge the dataset with itself based on the player's name
merged_data <- merge(x = Acquired_subset, y = Relinquished_subset, by.x = "Acquired", by.y = "Relinquished", suffixes = c("_end", "_begin"))

merged_data$Date_gap <- as.numeric(difftime(merged_data$Date_end, merged_data$Date_begin, units = "days"))

# Removing all Date_gaps less than 0
merged_data <- subset(merged_data, Date_gap > 0 )

# Filtering to ensure no duplicate Date_end for a given Date_begin
injury_duration <- merged_data %>% 
  group_by(Acquired, Date_end)  %>% 
  filter(Date_gap == min(Date_gap))

# Filtering to ensure no duplicate Date_end for a given Date_begin
injury_duration <- injury_duration %>% 
  group_by(Acquired, Date_begin)  %>% 
  filter(Date_gap == min(Date_gap))

# Reordering collumns of data
injury_duration <- injury_duration[, c("Acquired", "Date_begin", "Date_end", "Date_gap", "Team_begin", "Team_end", "Notes_begin", "Notes_end")]

# Rename column to Player
colnames(injury_duration)[colnames(injury_duration) == "Acquired"] <- 'Player'

Injury_duration_average <- mean(injury_duration$Date_gap)
Injury_duration_average

filtered_injury_duration <- subset(injury_duration, Date_gap <= 365 )
Injury_duration_average <- mean(filtered_injury_duration$Date_gap)
Injury_duration_average

# This graph is messed up because Patty Mills had an injury that lasted for 2976 days, which is not accurate to reality
ggplot(injury_duration, aes(x = Date_gap)) +
  geom_bar(fill = "skyblue", color = "black") +
  labs(title = "Frequency of Different Injury Durations",
       x = "Duration of injury",
       y = "Frequency") +
  theme_minimal()

# This graph is a zoomed in version of only the injuries that lasted less than a year, this should filter out the injuries that were not accurate.
ggplot(filtered_injury_duration, aes(x = Date_gap)) +
  geom_bar(fill = "skyblue", color = "black") +
  labs(title = 
  "Frequency of Different Injury Durations
  Excluding Injuries that Lasted more than a year (365 days)",
       x = "Duration of injury",
       y = "Frequency") +
  theme_minimal()
```

Based off of this data the average duration of an injury will be 11 days when you exclude the extereme injuries that lasted more than a year. We can see that the data is massively skewed towards shorter injuries. The reason for this being that many of the data points aren't injuries at all and instead are illness. Some instants of players being placed on IL don't have any detail of why. This makes it hard for us to filter out only injuries.\
One major issue with this data is it does not separate rest days and illness from serious injuries. We could filter out based on a certain duration of injury and label them as "serious". Labeling our data this way would give us a subset of data with supposedly "serious" injuries.

## Injury per Player data setup
```{r, message=FALSE}
library(dplyr)
library(lubridate)
#change injurydata date column to just the year instead of date format to help merge
changed_injurydata <- filtered_injurydata
changed_injurydata$Date <- as.Date(changed_injurydata$Date)
changed_injurydata$Date <- year(changed_injurydata$Date)

# merge datasets and combine player info and injured player
ageVSinjury <- inner_join(changed_injurydata, filtered_seasonsStatsdata, 
                         by = c("Relinquished" = "Player", "Date" = "Year"),
                         relationship = "many-to-many")

# make sure there are no repeats (distinct IDs)
ageVSinjury <- distinct(ageVSinjury,...1.x, .keep_all = TRUE)
```
## Injuries per Player (2010 - 2017)
```{r, message=FALSE}
# First look at age range
age_range <- summary(filtered_seasonsStatsdata$Age)
age_range 
# Age range is 19-40, lets split into groups by 3-4

# With this knowledge we can make a function to split up a dataset into a list with different age groups
SplitAges <- function(df) {
  age1 <- subset(df, Age <= 22)
  age2 <- subset(df, Age <= 26 & Age > 22)
  age3 <- subset(df, Age <= 30 & Age > 26)
  age4 <- subset(df, Age <= 35 & Age > 30)
  age5 <- subset(df, Age > 35)
  return(list(age1 = age1,age2 = age2,age3 = age3,age4 = age4,age5 = age5)) #Returns list of 5 lists of age groups
}
# Split age range by calling function
ageInjury_list <- SplitAges(ageVSinjury["Age"])
agePlayer_list <- SplitAges(filtered_seasonsStatsdata["Age"])

# Function to print/save age ratios for each group
AgeRatios <- function(list1,list2,PRINT) {
  if(PRINT == TRUE) {
    print(paste("19-22 years ratio:", nrow(list1[[1]]) / nrow(list2[[1]])))
    print(paste("23-26 years ratio:", nrow(list1[[2]]) / nrow(list2[[2]])))
    print(paste("27-30 years ratio:", nrow(list1[[3]]) / nrow(list2[[3]])))
    print(paste("31-35 years ratio:", nrow(list1[[4]]) / nrow(list2[[4]])))
    print(paste("36-40 years ratio:", nrow(list1[[5]]) / nrow(list2[[5]])))
  } else {
    ratio1 <- nrow(list1[[1]]) / nrow(list2[[1]])
    ratio2 <- nrow(list1[[2]]) / nrow(list2[[2]])
    ratio3 <- nrow(list1[[3]]) / nrow(list2[[3]])
    ratio4 <- nrow(list1[[4]]) / nrow(list2[[4]])
    ratio5 <- nrow(list1[[5]]) / nrow(list2[[5]])
    return(list("19-22 years" = ratio1,"23-26 years" = ratio2,"27-30 years" = ratio3,"31-35 years" = ratio4,"36-40 years" = ratio5))
  }
}

# Print out ratios of injured players vs all players for each age range
AgeRatios(ageInjury_list, agePlayer_list, TRUE) # PRINTS AGE RATIOS
listAgeRatios <- AgeRatios(ageInjury_list, agePlayer_list, FALSE) # SAVES AGE RATIOS AS LIST
```
## Graph for Injuries per player
```{r, message=FALSE}
library(ggplot2)
# Convert ratios to data frame
dfAgeRatio <- data.frame(Age_Group = names(listAgeRatios), Ratio = unlist(listAgeRatios), row.names = NULL)
dfAgeRatio
# Create graph to display Age vs Injury data
ggplot(dfAgeRatio, aes(x = Age_Group, y = Ratio, fill = Age_Group)) +
  geom_bar(stat = "identity") +
  labs(title = "Injuries per Player by Age Group in the NBA", x = "Age Group", y = "Injuries per Player") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

This graph was created to visualize the difference in injuries per age group. The injuries per player above 1 mean there is more occurrences of players getting injured then there is players in the league. So this means there are players getting injured multiple times or having reoccurring injuries. We also can see from the graph that the most injuries are occurring when players are young and/or just entering the NBA. This could be due to their young bodies still developing, they have not quite got to male physical bodily peak. This could also be due to the heavy workload in the NBA that they are not used to. We also see older players getting injured highly as expected. As you get older the more prone to injury you are as your body slows down. 

## Statistics for injury duration by age group
```{r, message=FALSE}
library(lubridate)
# Change date column to just year
changed_injuryduration <- injury_duration
changed_injuryduration$Date_begin <- as.Date(changed_injuryduration$Date_begin)
changed_injuryduration$Date_begin <- year(changed_injuryduration$Date_begin)

# Merge injury_duration and filtered_seasonsStatsdata
durationANDage <- inner_join(changed_injuryduration, filtered_seasonsStatsdata, 
                             by = c("Date_begin" = "Year", "Player" = "Player"),
                             relationship = "many-to-many")
# Split by age group
ageInjuryDuration_list <- SplitAges(durationANDage[c("Date_gap", "Age")])

# Statistics for each age group
statisticCalculations <- function(df, fun) {
  fun1 <- fun(df[[1]]$Date_gap)
  fun2 <- fun(df[[2]]$Date_gap)
  fun3 <- fun(df[[3]]$Date_gap)
  fun4 <- fun(df[[4]]$Date_gap)
  fun5 <- fun(df[[5]]$Date_gap)
  return(list("19-22 years" = fun1, "23-26 years" = fun2, "27-30 years" = fun3, "31-35 years" = fun4, "36-40" = fun5))
}
# Get lists for each age group for means, maxs, medians, sds
meanDurations <- statisticCalculations(ageInjuryDuration_list, mean)
maxDurations <- statisticCalculations(ageInjuryDuration_list, max)
medianDurations <- statisticCalculations(ageInjuryDuration_list, median)
sdDurations <- statisticCalculations(ageInjuryDuration_list, sd)
```
## Graph Injury duration by age group
```{r, message=FALSE}
#install.packages('cowplot')
library(cowplot)
library(ggplot2)
# Create datasets for the statistics for graph building
mean_df <- data.frame(Age_Group = names(meanDurations), Mean = unlist(meanDurations))
max_df <- data.frame(Age_Group = names(maxDurations), Max = unlist(maxDurations))
median_df <- data.frame(Age_Group = names(medianDurations), Median = unlist(medianDurations))
sd_df <- data.frame(Age_Group = names(sdDurations), SD = unlist(sdDurations))


statistic_barplot <- function(data, x, y, fill, title, y_label) {
  plot <- ggplot(data, aes_string(x = x, y = y, fill = fill)) +
    geom_bar(stat = "identity") +
    labs(title = title, x = "Age Group", y = y_label) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
          plot.title = element_text(size = 11),
          axis.title = element_text(size = 8),
          legend.text=element_text(size=8),
          legend.title = element_text(size = 10))
  return(plot)
}
# Create barcharts for Mean, Max, Median,  by age group
bar_means <- statistic_barplot(mean_df, "Age_Group", "Mean", "Age_Group", "Mean Injury Duration by Age Group", "Mean Injury Duration (Days)")
bar_maxs <- statistic_barplot(max_df, "Age_Group", "Max", "Age_Group", "Longest Injury Duration by Age Group", "Longest Injury Duration (Days)")
bar_medians <- statistic_barplot(median_df, "Age_Group", "Median", "Age_Group", "Median Injury Duration by Age Group", "Median Injury Duration (Days)")
bar_sd <- statistic_barplot(sd_df, "Age_Group", "SD", "Age_Group", "Standard Deviation Injury Duration by Age Group", "Standard Deviation Injury Duration (Days)")

plot_grid(bar_means, bar_medians, bar_maxs, bar_sd, ncol = 2, nrow = 2, align = "v")

```

We created these graphs to visualize some statistics of the players' injury duration by age group to see if we can find any trends. Looking at the 36-40 year age group we see it has the highest mean, lowest median, longest injury duration, and highest standard deviation. From this we can conclude the data is very spread out, more specifically skewed right. This means while on average there may be older players recovering quickly from injuries, there is also a significant proportion of players who experience longer recovery times. This trend is shown for all age groups which tells us in all age groups there are some injuries that are extremely long to recover from. This trend is mostly profound in the 36-40 year old age group.

## Look at injuries over time 2010 - 2022
```{r,message=FALSE}
library(ggplot2)
#change injurydata date column to just the year instead of date format to help merge
changed_injurydata <- filtered_injurydata
changed_injurydata$Date <- as.Date(changed_injurydata$Date)
changed_injurydata$Date <- year(changed_injurydata$Date)

# Prepare data (create dataset with each year and amount of injuries)
injuryOverTime <- data.frame(
  Year = unique(changed_injurydata$Date),
  Injuries = sapply(unique(changed_injurydata$Date), function(year) {
    sum(changed_injurydata$Date == year)
  })
)
# Exclude year 2023
injuryOverTime <- subset(injuryOverTime, Year != 2023)

# Convert Year column to factor with ordered levels
injuryOverTime$Year <- factor(injuryOverTime$Year, levels = unique(injuryOverTime$Year))

# Create a line plot using ggplot2
ggplot(injuryOverTime, aes(x = Year, y = Injuries, group = 1)) +
  geom_point() +
  geom_line() +  # Add points for each data point
  labs(x = "Year", y = "# of Injuries", title = "Injury Count Over Years") +
  scale_x_discrete(labels = as.character(injuryOverTime$Year))

```

We created a line and dot plot to see any trends between injuries and years. This helps us visualize if injury prevention has gotten better, worse, or stayed the same. We see an upward trend for the most part as we go through the years. We also see some spike ups and downs, let's dive into those further individually one at a time.

2020 was the year of the pandemic outbreak. This caused the season to be suspended and have restrictions on playing. This caused less games which means less oppurtunities to get injured. There was an effort to bring back nba games with the bubble inviting 22 teams to play basketball games through rigorous testing and caution. Keep in mind for later that some players and teams did not get involved with this bubble season. After this year we see an outbreak of injuries instead of sickness.

In 2021 there could be many causes for the injury outbreak. Some could be the abrupt start back to intense play and the shortened off season. The 2020-2021 season started December 22, 2020, so most of this season is played through 2021. The bubble from the 2019-2020 season had many factors on the injury outbreak in 2021. This caused some players who played in it to have a shortened off season and caused the players who didn't play, a long time off of intense activity. The bubble season was very condensed and could have overworked the players. Leading into a short off season did not give them enough time to recover and get ready for the next season. For the players who didn't play this might have been too long of time off, so when play started again the players might have overworked their body that stopped being used to this intense activity.

We see a good increase of injuries in 2017. This could be caused by the increased pace of the game. In the 2017-2018 there were many timeout changes made to the rulebook. The main focus of these changes were increasing the pace of the game. Instead of some 90 second timeouts and some 60 second timeouts, all timeouts will be 75 seconds flat. Through this and other changes to mandatory timeouts and stricter delay of game policies if you go over the time, players saw shorter breaks and more time in the game without a break, especially in the last few minutes of a game which sometimes can be the most intense on your body.

```{r}
#combining data from injury_data and player_data, replacing values in Acquired column if NA shows up in Acquired column with Relinquished value in row. Then changing name to equal Acquired.

#We are making a column to merge by in the future

combining_data <- injury_data %>% select(Acquired, Relinquished, Notes) %>% mutate(Acquired = if_else(is.na(Relinquished), Acquired, Relinquished)) %>% mutate(name = Acquired) 


#made a new column with name to align with player_data, but no distinction among names, so same name will show up multiple times

#below will merge the two dataframes by name

merged_playerData_combiningData <- merge(player_data,combining_data, by = "name")

#count of injuries per player
counting_number_injuries <- combining_data %>% count(Acquired) #%>% as.numeric(n) 

ggplot(counting_number_injuries, aes(x = n)) +
  geom_histogram(binwidth = 1,) +
  labs(title = "Frequency of Injuries", x = "number of injuries", y = "Frequency")
```
This shows that the number of injuries is right-skewed, so little to none have had 75 injuries or up. The number of injuries at the beginning is very high considering the fact that the NBA is a very competitive environment. All that competitivity leads to more physical interactions between players. 


```{r}
#counts the number of injuries 
#counting_number_of_injuries <- counting_number_injuries %>% rename(number_of_injuries = n) %>% count(number_of_injuries) %>% rename(people_injured = number_of_injuries) %>% rename(count_of_injuries = n)

#print(counting_number_of_injuries)

merged_playerData_combiningData <- merge(player_data,combining_data, by = "name")




# shows number of injuries based on position
injuries_by_position <- merged_playerData_combiningData %>% count(position) %>% rename(number_of_injuries = n)

library(ggplot2)

#creates bar graph of injuries by position

ggplot(injuries_by_position, aes(x = position, y = number_of_injuries)) +
  geom_bar(stat = "identity") +
  labs(x = "Position Injuries", y = "Frequency", title = "Frequency of Injuries by Position")
```
This shows that the forward and guard get the most amount of injuries. The type of injury is uncertain considering we haven't differentiated between IL (can't play due to injury, but takes up bench spot) and IR (can't play due to injury, but doesn't take up a bench spot). 

```{r}
##IL means they won't take up a bench spot
##IR means they'll be taking up a bench spot

##would certain positions get put on IL instead of IR because of position played?

library(stringr)
IL <- "placed on IL"
IR <- "placed on IR"

#below gives injuries that start with "placed on IL"

IL_position_injuries <- merged_playerData_combiningData %>% filter(str_starts(Notes, IL)) %>% select(position, Notes) %>% count(position) %>% rename(IL_injuries = n)

#below gives something that starts with IR and counts the position of each person

IR_position_injuries <- merged_playerData_combiningData %>% filter(str_starts(Notes, IR)) %>% select(position, Notes) %>% count(position) %>% rename(IR_injuries = n)
#print(IR_position_injuries)

#above from counting position for people placed on IR shows guard and forward are placed on it the most, could possibly be due to aggressiveness of position needed, however F-G, and Guard forward are the lowest

ggplot(IR_position_injuries, aes(x = position, y = IR_injuries)) +
  geom_bar(stat = "identity") +
  labs(x = "Position", y = "Frequency", title = "IR position injuries")

##how many IL happen per position, but by same position how many IR happens?

##IL means a more serious injury which could then impact that persons life. 
##Would they put someone on IL if they had a certain position compared to another position?

ggplot(IL_position_injuries, aes(x = position, y = IL_injuries)) +
  geom_bar(stat = "identity") +
  labs(x = "Position Injuries", y = "Frequency", title = "IL Position Injuries")


#merging IR and IL injuries in order to compare in bar graph
merging_IR_IL <- merge(IL_position_injuries, IR_position_injuries, by = "position")
#print(merging_IR_IL)


#bar graph below shows how much more likely you are to get put on IL when at that position.
#Doesn't account for type of injury, aggressiveness needed at said position and how they affect the team chemistry

ggplot(merging_IR_IL, aes(x = position, y = IL_injuries / IR_injuries)) +
  geom_bar(stat = "identity") +
  labs(x = "Position", y = "IL Injuries per IR Injuries", title = "IL to IR Injury constant Based on Position")


library(tidyr)


#below gets the number of players injured
injured_sample <- merged_playerData_combiningData %>% distinct(name) %>% nrow()

#below gets the total number of players
overall_sample_size <- player_data %>% distinct(name) %>% nrow()

#below gets the injury at least once per total players
print(injured_sample / overall_sample_size) * 100

#above shows we got a 46.5% chance a player will get injured at least once, while NLH says the amount injured at least once is a 46.8% chance
 

```
The first graph shows how how many IR position injuries there are including all years from 1947 to 2018.
The second graph shows how many IL position injuries including 2010.
The third graph shows how much more likely you are to get an IL injury, calculated using IL injuries per IR injuries.
The fourth number shows the likelihood of one injury occurring, in this case 46.5%, NLH got a number of 46.8%.

```{r}
combining_data_any_year <- injury_data %>% select(Acquired, Relinquished, Notes) %>% mutate(Acquired = if_else(is.na(Relinquished), Acquired, Relinquished)) %>% mutate(name = Acquired) 

merged_playerData_combiningData_any_player <- merge(player_data, combining_data_any_year, by = "name")


#below gives IL injuries for any year

IL_position_injuries_any_year <- merged_playerData_combiningData_any_player %>% filter(str_starts(Notes, IL)) %>% select(position, Notes) %>% count(position) %>% rename(IL_injuries = n)

#below gives IR's for any amount of years

IR_position_injuries_any_year <- merged_playerData_combiningData_any_player %>% filter(str_starts(Notes, IR)) %>% select(position, Notes) %>% count(position) %>% rename(IR_injuries = n)

#for any year
IR_position_injuries_any <- merged_playerData_combiningData %>% filter(str_starts(Notes, IR)) %>% select(position, Notes) %>% count(position) %>% rename(IR_injuries = n)


IR_position_injuries_any_year <- merged_playerData_combiningData_any_player %>% filter(str_starts(Notes, IR)) %>% select(position, Notes) %>% count(position) %>% rename(IR_injuries = n)

#how many IR's for any year?
ggplot(IR_position_injuries_any_year, aes(x = position, y = IR_injuries)) +
  geom_bar(stat = "identity") +
  labs(x = "Position", y = "Frequency", title = "IR position injuries for any year")

```



This shows that the IR position injuries for any year still has forward and guard as the most injured player positions. This could mean a tiny IL injury weakens that individuals body, then it gets hurt again and the injury becomes more severe than last time.

# Conclusions

Considering the NLM article is a recognized and peer-reviewed article, in regards to the probability of an injury from 1947 to 2018 is accurate, but off by NLM's percentage by 0.3%. The NLM's link below.

https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6107769/#:~:text=For%20professional%20athletes%2C%20injuries%20can,negative%20consequences%20for%20many%20stakeholders. 

A player tends to get hurt a couple times, but after about twenty five sustained injuries, they tend to not get injured as much. 

IR injuries don't happen as often as IL injuries. The number of IL injuries is in the thousands per position, but for IR injuries it tend to be in the hundreds. This absence of IR injuries could be due to the classification of IR injuries being changed after the year 2006.

IL injuries are more than 15 times likely to happen to you compard to IR injuries.
